Native advertising as a content marketing strategy

### 2024-09-10 analysis/paper/paper.Rmd
los

The concentration of so-called hybrid forms of media content in the contemporary online environment is becoming increasingly intense (Balasubramanian, 1994; Macnamara, 2014; Taiminen, Luoma-aho & Tolvanen, 2015). Among the most prominent forms are sponsored content (Tutaj & Reijmersdal, 2012), content marketing (Pulizzi, 2014; Pulizzi & Piper, 2023), brand journalism (Cole & Greer, 2013), and native advertising (Verčič & Tkalac Verčič, 2016). For most researchers, the latter term will serve as an umbrella term that attempts to simplify the classification of hybrid forms, as native advertising by definition implies a paid advertisement that resembles editorial content, aiming to attract the target audience with useful content while simultaneously promoting the brand, values, and products (Cole & Greer, 2013; Howe & Teufel, 2014; Wojdynski & Evans, 2016). On the other hand, the research by Taiminen et al. (2015) showed that most public relations and marketing professionals preferred to use the term “content marketing” when discussing commercial hybrid forms of online content. In addition to the obvious terminological confusion, three other problematic areas of native advertising need to be highlighted:

@@

>	Legal and ethical framework: How is native advertising legally defined, and what is the practice of media houses in labelling such advertisements (Trbušić, Maleš & Šikić, 2024).

The theoretical basis of this paper is the thesis that native advertising is “an extension of content marketing and provides marketers with a platform to place content in front of a wider audience they wouldn't ordinarily reach” (Crook, 2022). Although there are authors who highlight certain criteria by which content marketing differs from native advertising, which will be described in more detail later in the paper, the ultimate goal is common: to drive profitable customer action. The profitability of such market communication for organizations and brands is demonstrated by recent metrics. Figures show that 47% of marketing professionals globally believe in the effectiveness of this type of advertising (Kloot, 2022), and that in 2023, spending on native advertising in the United States increased by 12% compared to the previous year (eMarketer, 2023). Thus, with more than 97 billion US dollars invested, native advertising holds the largest share (59.7%) of the total expenditure for all types of advertising (eMarketer, 2023).

@@

Charmaine Du Plessis (2015a) defines content marketing as sharing brand content via owned media to gain earned media, which clearly implies that it is not a type of media content that is paid for, as (native) advertising certainly is. In this context, Du Plessis employs the classic PESO model of media content in public relations (Dietrich, 2014), which distinguishes paid media (advertising or other sponsored content); earned media (publicity generated from media pitches or news releases); shared media (content shared and re-shared on social media); and owned media (such as company’s website, newsletter, or email communications). The order in the PESO model has long reflected the priorities of organizations in media strategy planning. However, recent research (Macnamara, Lwin, Adi & Zerfass, 2016) indicates a shift in the order in practice, with the dominance shifting to the SOEP model instead of the classic PESO model. The reasons for this are manifold, but the most indicative is that owned and earned media content are trusted more than paid media, while shared media (i.e., “recommendations from people I know” and “consumer opinions posted online”) are the most trusted information sources of all (Nielsen, 2015).
Like Du Plessis, other authors (Handley & Chapman, 2011; Halvorson & Rach, 2012) also differentiate content marketing from other forms of hybrid content exclusively based on the distribution channel (owned media). However, there are authors who leave much more room for interpretation, not strictly confining it to owned media that later generates earned media (publicity). Their definitions are more focused on the strategy, intent, and ultimate goal of content marketing, rather than on the tactics and type of media content within the PESO framework. For instance, they highlight that content marketing is “a strategic marketing approach focused on creating and distributing valuable, relevant, and consistent content” (Pulizzi & Piper, 2023; CMI, 2024), which is disseminated on various platforms such as newspapers, magazines, TV, or radio channels (Holliman & Rowley, 2014).

The distribution of valuable, relevant, and consistent content is also at the core of native advertising (Campbell & Marks, 2015; Kim, 2017), which is especially evident in so-called branded or native content. According to the Interactive Advertising Bureau (IAB), there are three fundamental forms of native advertising: In-Feed/In-Content native advertising, Content Recommendation Ads, and Branded/native content. The latter refers to “paid content by a brand that is published in the same format as full-fledged editorial content on the publisher's website, usually in collaboration with the publisher's own teams” (IAB, 2019, p. 14). This form of native advertising will be the focus of the research part of this paper for two main reasons: it conceptually correlates with the forms and tactics of content marketing, and it is the only form among the mentioned types of native advertising that can be approached analytically.

Types and tactics of content marketing typically include blogs, infographics, video content, photos, webinars, podcasts and social media (Halligan & Shah, 2010; Handley & Chapman, 2011; Bloomstein, 2012; Wuebben, 2012; Vibrant Publishers, 2020), but also paid ads (Baker, 2024), such as sponsored advertising options on social media, affiliate marketing, remarketing, native ads etc. (Jefferson & Tanton, 2015). It is yet another proof of the inconsistent interpretation of content marketing as something that is disseminated exclusively via owned media. It is more appropriate to consider content marketing as a set of tactics that, among other things, can drive traffic back to the website of some company or product (owned media). It is also noticeable that some authors of definitions avoid identifying content marketing with product advertising which aims at the purchase of a product or service (Patrutiu Baltes, 2015), seeking to further differentiate it from other related marketing tactics. The fact is that content marketing is part of an inbound marketing strategy, which, unlike the outbound strategy, is focused on “attracting valuable consumers (potential, existing, or aspirational) who choose to interact with a particular company that provides them with something useful” (Opreana & Vinerean, 2015, p. 29). However, it is misguided to claim that the ultimate goal of any marketing communication, including content marketing, is not sales or gaining profit (Rowley, 2008; Pulizzi & Piper, 2023). Furthermore, both academics and practitioners agree that content marketing relies on a consumer-centric strategy (Ho, Pang & Choy, 2020). In conclusion, regardless of the type of marketing communication, whether it is content marketing, native advertising, or brand journalism - although one can discuss nuanced tactics and approaches to attract consumers into the marketing funnel and later motivate them for retention and advocacy phases - profitable customer action remains the intrinsic motive of every organization or brand.

Previous research does not encompass the analysis of native advertising as part of content marketing; rather, these concepts have been the focus of separate studies. Thus, insights in the context of native advertising relate to the analysis of readers' cognitive perception of native ads and their persuasive elements (Wojdynski & Evans, 2016), analysis of communication strategies in native ads (Wang & Li, 2017), perception of the organization or the sponsor of native ads and its socially responsible activities (Jung & Heo, 2018; Krouwer, Poels & Paulussen, 2019; Beckert, Koch, Viererbl, Denner, & Peter, 2020; Wu & Overton, 2021), perception of the credibility of online portals that publish native ads (Howe & Teufel, 2014), and analysis of the basic features of native ads based on the most-read selected Croatian online portals (Trbušić et al., 2024). Research in the field of content marketing primarily reflects valuable theoretical contributions, such as conceptual definitions, classification of types and forms of content marketing, identification of strategies and tactics of content marketing in the digital environment (Koiso-Kanttila, 2004; Rowley, 2008; Handley & Chapman, 2011; Bloomstein, 2012; Wuebben, 2012; Pulizzi, 2014; Jefferson & Tanton, 2015; Vinerean, 2017; Pulizzi & Piper, 2023), and the definition of six fundamental elements of content marketing which should be considered by marketers (Du Plessis, 2015a; Du Plessis, 2015b). Empirically, content marketing has been analysed as part of B2B strategies on digital channels (Holliman & Rowley, 2014) or in the context of content creation practices in successful companies (Du Plessis, 2015c; Ho et al., 2020). Additionally, studies have addressed the causal linkage between content marketing and online consumer behaviour (Du Plessis, 2022) and content marketing as part of business strategy in the luxury industry (Rios, 2016).

@@

The sample for the study consists of native advertisements published over a six-month period (December 2021 – May 2022) on the six most-read Croatian internet portals (index.hr, jutarnji.hr, vecernji.hr, 24sata.hr, telegram.hr, slobodnadalmacija.hr), which were identified as the most influential online media in the Croatian digital space based on visitor criteria (Peruško, 2023). Following a machine-based pre-analysis that included all posts on Croatian internet portals during the observed period, totalling over 1.5 million posts on 5,104 web pages, a representative sample was selected based on visit frequency, specifically from the largest Croatian internet portals. A machine search was conducted using keywords and phrases in the content or headlines. The keywords used to identify native content included phrases such as “paid content”, “sponsored content”, “native”, “content provided by”, “content created by”, “powered by” and “tnative”, as native content should be clearly identified as such in the media. After filtering all articles from the six-month period that matched these keywords, a total of 811 articles were identified, from which non-native articles, primarily sponsored or PR content and other irrelevant materials, were manually excluded. The final sample for the study consisted of 543 native advertisements across the six internet portals noted for their influence and readership in Croatia. Prior to determining the research sample, it was confirmed that all analysed internet portals offer native advertising services through a review of their publicly available marketing, advertising, and public relations pricing lists.

@@

For the purposes of this study, two general categories were observed, in line with content marketing types and typologies: headline characteristics and content characteristics of native advertisements. The focus was on categorizing headline types, their connection with the advertiser's brand, and identifying the type of clickbait headline. The content analysis of native articles focused on the type of visual equipment (photo, video, infographic, graphics, combinations of these) and its relationship with the advertiser's visual identity, as well as the representation of different types of sources/statements (source from the organization/company, brand ambassador, influencer/celebrity, direct user of the product or service, multiple sources) in native ads. The study also examined whether certain types of native ad headlines were more prevalent or characteristic of different industries. Similarly, the types of statements found in the analysed native ads were compared and observed concerning the industry type and the reach of each ad.

@@

In determining clickbait headlines, the categorization provided by Chen et al. (2015) was used: vague pronouns, uncertainty, reverse narrative style, emphasis on emotions, and the use of numbers. The study revealed that the most prevalent clickbait headlines were those characterized by uncertainty (44.38%). Multiple combinations of styles within headlines were observed in 13.81% of the ads, while 20.99% of the headlines exhibited no clickbait characteristics at all. Headlines dominated by numerical content appeared in 9.39% of cases, while 5.89% of headlines emphasized emotional appeal. Headlines using vague pronouns were noted in 4.97% of cases, and the reverse narrative style was registered in 0.55% of the ads.

@@

The theoretical basis of this paper was the thesis that native advertising is “an extension of content marketing and provides marketers with a platform to place content in front of a wider audience they wouldn't ordinarily reach” (Crook, 2022). Although there are authors who highlight certain criteria by which content marketing differs from native advertising, the ultimate goal is common: to drive profitable customer action. Based on the findings of this research, several conclusions can be drawn. The analysis of 543 native ads published on six Croatian online portals reveals a predominant use of clickbait headlines, with nearly 80% of headlines containing at least one clickbait characteristic. The most frequently employed technique is the style of uncertainty, which significantly contributes to audience engagement, as evidenced by its high average and total reach. This indicates that uncertainty-driven headlines are particularly effective in attracting readers. This is evidence that advertisers, through their content, aim to promote a “culture of fear”, a concept typically explored by sociologists (Maskalan, 2014; Furedi, 1997) in the context of political actions, but which is also applicable in the advertising industry. This is particularly due to the fact that fear/insecurity appeals can encourage people to change their attitudes and purchasing preferences (Foxall et al., 2007; Kesić, 2003). Martin Lindstrom (2008) anticipated that contemporary advertising would increasingly be based on somatic markers triggered by fear and a sense of insecurity. In other words, advertisers will try to convince us that by not purchasing their product, we become insecure, dissatisfied, and unfree, which is evident in this analysis as well.

>>>>

The concentration of so-called hybrid forms of media content in the contemporary online environment is becoming increasingly intense [@balasubramanian1994; @macnamara2014; @taiminen2015]. Among the most prominent forms are sponsored content [@tutaj2012], content marketing [@pulizzi2014; @pulizzi2023], brand journalism [@cole2013], and native advertising [@vercic2016]. For most researchers, the latter term will serve as an umbrella term that attempts to simplify the classification of hybrid forms, as native advertising by definition implies a paid advertisement that resembles editorial content, aiming to attract the target audience with useful content while simultaneously promoting the brand, values, and products [@cole2013; @howe2014; @wojdynski2016]. On the other hand, the research by [@taiminen2015] showed that most public relations and marketing professionals preferred to use the term “content marketing” when discussing commercial hybrid forms of online content. In addition to the obvious terminological confusion, three other problematic areas of native advertising need to be highlighted:

@@

>	Legal and ethical framework: How is native advertising legally defined, and what is the practice of media houses in labelling such advertisements [@trbusic2024]?

The theoretical basis of this paper is the thesis that native advertising is “an extension of content marketing and provides marketers with a platform to place content in front of a wider audience they wouldn't ordinarily reach” [@crook2022]. Although there are authors who highlight certain criteria by which content marketing differs from native advertising, which will be described in more detail later in the paper, the ultimate goal is common: to drive profitable customer action. The profitability of such market communication for organizations and brands is demonstrated by recent metrics. Figures show that 47% of marketing professionals globally believe in the effectiveness of this type of advertising [@kloot2022], and that in 2023, spending on native advertising in the United States increased by 12% compared to the previous year [@emarketer2023]. Thus, with more than 97 billion US dollars invested, native advertising holds the largest share (59.7%) of the total expenditure for all types of advertising [@emarketer2023].

@@

[@duplessis2015a] defines content marketing as sharing brand content via owned media to gain earned media, which clearly implies that it is not a type of media content that is paid for, as (native) advertising certainly is. In this context, Du Plessis employs the classic PESO model of media content in public relations [@dietrich2014], which distinguishes paid media (advertising or other sponsored content); earned media (publicity generated from media pitches or news releases); shared media (content shared and re-shared on social media); and owned media (such as company’s website, newsletter, or email communications). The order in the PESO model has long reflected the priorities of organizations in media strategy planning. However, recent research [@macnamara2016] indicates a shift in the order in practice, with the dominance shifting to the SOEP model instead of the classic PESO model. The reasons for this are manifold, but the most indicative is that owned and earned media content are trusted more than paid media, while shared media (i.e., “recommendations from people I know” and “consumer opinions posted online”) are the most trusted information sources of all [@nielsen2015].
Like Du Plessis, other authors [@handley2011; @halvorson2012] also differentiate content marketing from other forms of hybrid content exclusively based on the distribution channel (owned media). However, there are authors who leave much more room for interpretation, not strictly confining it to owned media that later generates earned media (publicity). Their definitions are more focused on the strategy, intent, and ultimate goal of content marketing, rather than on the tactics and type of media content within the PESO framework. For instance, they highlight that content marketing is “a strategic marketing approach focused on creating and distributing valuable, relevant, and consistent content” [@pulizzi2023; @cmi2023], which is disseminated on various platforms such as newspapers, magazines, TV, or radio channels [@holliman2014].

The distribution of valuable, relevant, and consistent content is also at the core of native advertising [@campbell2015; @kim2017], which is especially evident in so-called branded or native content. According to the Interactive Advertising Bureau (IAB), there are three fundamental forms of native advertising: In-Feed/In-Content native advertising, Content Recommendation Ads, and Branded/native content. The latter refers to “paid content by a brand that is published in the same format as full-fledged editorial content on the publisher's website, usually in collaboration with the publisher's own teams” [@iab2019]. This form of native advertising will be the focus of the research part of this paper for two main reasons: it conceptually correlates with the forms and tactics of content marketing, and it is the only form among the mentioned types of native advertising that can be approached analytically.

Types and tactics of content marketing typically include blogs, infographics, video content, photos, webinars, podcasts and social media [@halligan2010; @handley2011; @bloomstein2012; @wuebben2011; @vibrant2020], but also paid ads [@baker2024], such as sponsored advertising options on social media, affiliate marketing, remarketing, native ads etc. [@jefferson2015]. It is yet another proof of the inconsistent interpretation of content marketing as something that is disseminated exclusively via owned media. It is more appropriate to consider content marketing as a set of tactics that, among other things, can drive traffic back to the website of some company or product (owned media). It is also noticeable that some authors of definitions avoid identifying content marketing with product advertising which aims at the purchase of a product or service [@patrutiu2015], seeking to further differentiate it from other related marketing tactics. The fact is that content marketing is part of an inbound marketing strategy, which, unlike the outbound strategy, is focused on “attracting valuable consumers (potential, existing, or aspirational) who choose to interact with a particular company that provides them with something useful” [@opreana2015]. However, it is misguided to claim that the ultimate goal of any marketing communication, including content marketing, is not sales or gaining profit [@rowley2008; @pulizzi2023]. Furthermore, both academics and practitioners agree that content marketing relies on a consumer-centric strategy [@ho2020]. In conclusion, regardless of the type of marketing communication, whether it is content marketing, native advertising, or brand journalism - although one can discuss nuanced tactics and approaches to attract consumers into the marketing funnel and later motivate them for retention and advocacy phases - profitable customer action remains the intrinsic motive of every organization or brand.

Previous research does not encompass the analysis of native advertising as part of content marketing; rather, these concepts have been the focus of separate studies. Thus, insights in the context of native advertising relate to the analysis of readers' cognitive perception of native ads and their persuasive elements [@wojdynski2016], analysis of communication strategies in native ads [@wang2017], perception of the organization or the sponsor of native ads and its socially responsible activities [@jung2018; @krouwer2020; @beckert2020; @wu2021], perception of the credibility of online portals that publish native ads [@howe2014], and analysis of the basic features of native ads based on the most-read selected Croatian online portals [@trbusic2024]. Research in the field of content marketing primarily reflects valuable theoretical contributions, such as conceptual definitions, classification of types and forms of content marketing, identification of strategies and tactics of content marketing in the digital environment [@koiso2004; @rowley2008; @handley2011; @bloomstein2012; @wuebben2011; @pulizzi2014; @jefferson2015; @vinerean2017; @pulizzi2023], and the definition of six fundamental elements of content marketing which should be considered by marketers [@duplessis2015a; @duplessis2015b]. Empirically, content marketing has been analysed as part of B2B strategies on digital channels [@holliman2014] or in the context of content creation practices in successful companies [@duplessis2015c; @ho2020]. Additionally, studies have addressed the causal linkage between content marketing and online consumer behaviour [@duplessis2022] and content marketing as part of business strategy in the luxury industry [@rios2016].

@@

The data used in the study consists of native advertisements published over a six-month period (December 2021 – May 2022) on the six most-read Croatian internet portals (index.hr, jutarnji.hr, vecernji.hr, 24sata.hr, telegram.hr, slobodnadalmacija.hr) [@perusko2023]. The data was sampled based on the combination of machine and human filtering of the population of posts on Croatian internet portals during the observed period, totalling over 1.5 million posts on 5,104 web pages. The machine search was conducted based on selected keywords and phrases (“paid content”, “sponsored content”, “native”, “content provided by”, “content created by”, “powered by” and “tnative”) in the content or headlines of the article. This procedure resulted in 811 articles. After the thorough human filtering, irrelevant articles were manually excluded and the final empirical sample consisted of 543 native articles. Prior to determining the research sample, it was confirmed that all analysed internet portals offer native advertising services through a review of their publicly available marketing, advertising, and public relations pricing lists.

@@

For the purposes of this study, two general categories were selected for the analysis, in line with content marketing types and typologies: headline characteristics and content characteristics of native advertisements. The focus was on categorizing headline types, their connection with the advertiser's brand, and identifying the type of clickbait headline. The content analysis of native articles focused on the type of visual equipment (photo, video, infographic, graphics, combinations of these) and its relationship with the advertiser's visual identity, as well as the representation of different types of sources/statements (source from the organization/company, brand ambassador, influencer/celebrity, direct user of the product or service, multiple sources) in native ads. The study also examined whether certain types of native ad headlines were more prevalent or characteristic of different industries. Similarly, the types of statements found in the analysed native ads were compared and observed concerning the industry type and the reach of each ad.

@@

In determining clickbait headlines, the categorization provided by [@chen2015] was used: vague pronouns, uncertainty, reverse narrative style, emphasis on emotions, and the use of numbers. The study revealed that the most prevalent clickbait headlines were those characterized by uncertainty (44.38%). Multiple combinations of styles within headlines were observed in 13.81% of the ads, while 20.99% of the headlines exhibited no clickbait characteristics at all. Headlines dominated by numerical content appeared in 9.39% of cases, while 5.89% of headlines emphasized emotional appeal. Headlines using vague pronouns were noted in 4.97% of cases, and the reverse narrative style was registered in 0.55% of the ads.

@@

The theoretical basis of this paper was the thesis that native advertising is “an extension of content marketing and provides marketers with a platform to place content in front of a wider audience they wouldn't ordinarily reach” [@crook2022]. Although there are authors who highlight certain criteria by which content marketing differs from native advertising, the ultimate goal is common: to drive profitable customer action. Based on the findings of this research, several conclusions can be drawn. The analysis of 543 native ads published on six Croatian online portals reveals a predominant use of clickbait headlines, with nearly 80% of headlines containing at least one clickbait characteristic. The most frequently employed technique is the style of uncertainty, which significantly contributes to audience engagement, as evidenced by its high average and total reach. This indicates that uncertainty-driven headlines are particularly effective in attracting readers. This is evidence that advertisers, through their content, aim to promote a “culture of fear”, a concept typically explored by sociologists [@maskalan2014; @furedi1997] in the context of political actions, but which is also applicable in the advertising industry. This is particularly due to the fact that fear/insecurity appeals can encourage people to change their attitudes and purchasing preferences [@foxall2007; @kesic2003]. Martin Lindstrom [@lindstrom2008] anticipated that contemporary advertising would increasingly be based on somatic markers triggered by fear and a sense of insecurity. In other words, advertisers will try to convince us that by not purchasing their product, we become insecure, dissatisfied, and unfree, which is evident in this analysis as well.
